Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2020-01-09 Manuscript/manuscript.Rmd
edits

We analyzed publically available replicate Hi-C data sets from the GM12878 and K562 cell lines (Supplementary Table 1). Topologically associating domains (TADs) were subsequently called using the ARROWHEAD pipeline [@durand2016juicer] at five data resolutions: 5 kb 10 kb, 25 kb, 50 kb, and 100 kb. After removing TADs that exceeded 2 md in length (biologically plausible maximum TAD size [@dixon2012topological; @sofueva2013cohesin]), we obtained a total of 16133, 14824, 8081, 3560, and 919 unique TAD boundaries for the GM12878 cell line, across each of the five resolutions respectively (Table 1). Similar numbers of TAD boundaries were found for the K562 cell line. To account for the variability in boundary location due to varying resolution, we developed a resolution-specific approach toward constructing TAD versus non-TAD boundary regions on the linear genome (See Methods), labeled as a response vector **Y** (Figure 1A). TAD boundary regions were defined as regions on the linear genome where a genomic bin overlapped with a flanked TAD boundary ($Y = 1$), while non-TAD boundary regions were defined as the regions that did not overlap with a flanked TAD boundary ($Y = 0$). On average, we found that only $18.0 \pm 4.6 \%$ of TAD boundary regions overlapped between both cell lines. Given the cell type specificity of TAD boundaries, we opted to build cell line specific models for predicting TAD boundary regions. The results presented are for the GM12878 cell line, with K562 results presented in the Supplementary material. In summary, we developed a framework enabling resolution-specific prediction of TAD boundaries, with resolution-specific bins defining the response variable.

| Resolution | Total number of called TADs | Total number of unique TAD boundaries | Average Class Imbalance ($\pm$ SD) |
|:----------:|:---------------------------:|:-------------------------------------:|:----------------------------------:|
|5 kb        |8043 (4750)                  | 16133 (9314)                          | 0.06 $\pm$ 0.02 (0.04 $\pm$ 0.02)  |
|10 kb	     |7650 (5796) 	               | 14824 (10881)                         | 0.10 $\pm$ 0.01 (0.08 $\pm$ 0.02)  |
|25 kb	     |4208 (3675)                  | 8081 (6923)                           | 0.14 $\pm$ 0.01 (0.12 $\pm$ 0.01)  |   
|50 kb	     |1772 (1617)	                 | 3560	(3197)                           | 0.12 $\pm$ 0.01 (0.11 $\pm$ 0.01)  |
|100 kb	     |444	(399)                    | 919	(831)                            | 0.06 $\pm$ 0.01 (0.06 $\pm$ 0.01)  |    
Table: Domain data and class imbalance summaries across resolutions for GM12878 (K562).

>>>>

We analyzed publically available replicate Hi-C data sets from the GM12878 and K562 cell lines (Supplementary Table 1). Topologically associating domains (TADs) were subsequently called using the ARROWHEAD pipeline [@durand2016juicer] at five data resolutions: 5 kb 10 kb, 25 kb, 50 kb, and 100 kb. <!--After removing TADs that exceeded 2 md in length (biologically plausible maximum TAD size [@dixon2012topological; @sofueva2013cohesin]),--> We obtained a total of 16133, 14824, 8081, 3560, and 919 unique TAD boundaries for the GM12878 cell line, across each of the five resolutions respectively (Table 1). Similar numbers of TAD boundaries were found for the K562 cell line. To account for the variability in boundary location due to varying resolution, we developed a resolution-specific approach toward constructing TAD versus non-TAD boundary regions on the linear genome (See Methods), labeled as a response vector **Y** (Figure 1A). TAD boundary regions were defined as regions on the linear genome where a genomic bin overlapped with a flanked TAD boundary ($Y = 1$), while non-TAD boundary regions were defined as the regions that did not overlap with a flanked TAD boundary ($Y = 0$). On average, we found that only $18.0 \pm 4.6 \%$ of TAD boundary regions overlapped between both cell lines. Given the cell type specificity of TAD boundaries, we opted to build cell line specific models for predicting TAD boundary regions. The results presented are for the GM12878 cell line, with K562 results presented in the Supplementary material. In summary, we developed a framework enabling resolution-specific prediction of TAD boundaries, with resolution-specific bins defining the response variable.

| Resolution | Total number of called TADs | Total number of unique TAD boundaries | Total number of Genomic Bins | Class Imbalance |
|:----------:|:---------------------------:|:-------------------------------------:|:----------------------------:|:---------------:|
|5 kb        |8052                         | 15468                                 | 535363                       | 0.03            |
|10 kb	     |7676 	                       | 14253                                 | 267682                       | 0.05            |
|25 kb	     |4670                         | 8363                                  | 107073                       | 0.08            |
|50 kb	     |2349	                       | 4224                                  | 53537                        | 0.08            |
|100 kb	     |1031                         | 1883                                  | 26768                        | 0.07            |
Table: Domain data and class imbalance summaries across resolutions for GM12878.
